---
title             : "Do members of Congress have an increasing level of influence on the outcome of federal agency financial rulemaking?"
shorttitle        : "Have federal agencies become more partisan over time?"
author: 
  - name          : "Samiha Bhushan"
    affiliation   : "1"
    corresponding : no   # Define only one corresponding author
    email         : "bhushan3@wisc.edu"
affiliation:
  - id            : "1"
    institution   : "University of Wisconsin-Madison"
authornote: |
  Department of Political Science, University of Wisconsin-Madison
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
editor_options: 
  chunk_output_type: console
abstract: |
  empty
  
keywords          : "keywords"
wordcount         : "X"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
#output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```
#Theory
I will attempt to answer two questions: first, have federal agencies become increasingly receptive to the demands of members of Congress? Second, are more independent agencies less likely to be influenced by members of Congress? 

I use data from a previous study conducted by Jennifer Selin. Selin's data disseminates various structural features of agencies, such as if the agency is involved within an executive department, if the agency places a conflict of interest restriction on its members, if they provide some avenue for congressional input in the process it nominates its members, and if the agency is a commission or has a board of directors, whether the statute limits the number of members who may serve from the same party. 

With a Bayesian latent variable model, she estimates the level of independence for 321 federal agencies, where the higher the number, the more indendent the angency is in its rulemaking process. Extrapolating data from Selin, we find that financial agencies are generally rated higher in terms of level of independence, where the Consumer Financial Protection Bureau (CFPB) is rated one of the highest at 4.1 and the Federal Trade Commission (FTC) is 2.269. The CFPB is rated so highly because of its 5-year term limit for members, commissioners, and agency heads, additional clauses in the statute that enable members to be removed due to "neglect of duty", and a clause that enables the CFPB to fund a substantial amount of its funding from independent sources.  

# Methods
I will gather data from 40 rules issued by financial agencies, including but not limited to the Consumer Financial Protection Bureau, the Treasury Department, and the Office of the Comptroller of the Currency. I focus on public comments issued by members of Congress and focus on whether their demands were ultimately fulfilled in the final rule. 

The research design additionally combines data from a previous study conducted by Jennifer L. Selin, who determined an estimate of the level of structural independence for several agencies during the Obama administration. My study will also include rules during the subsequent Trump administration, wherein the Senate majority switched to the Republican party. I will compare and contrast rules from both eras to determine if influence from political parties have grown over time, including determining if there is an increased presence of politicians commenting on policies. I will also compare the estimates of agency independence as calculated by Selin to level of success with regard to demands made by members of Congress. 

I will also use textual analysis in addition to the hand coding to see if elected officials use more partisan rhetoric over time. In this study, “partisan rhetoric” refers to an individual making positive assertions about their own party or negative assertions about the opposing party. I will also examine whether the demands of elected officials that are the same party as the President are increasingly met over those of the opposing party. 

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results

# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
