---
title             : "Do members of Congress have an increasing level of influence on the outcome of federal agency financial rulemaking?"
shorttitle        : "Have federal agencies become more partisan over time?"
author: 
  - name          : "Samiha Bhushan"
    affiliation   : "1"
    corresponding : no   # Define only one corresponding author
    email         : "bhushan3@wisc.edu"
affiliation:
  - id            : "1"
    institution   : "University of Wisconsin-Madison"
authornote: |
  Department of Political Science, University of Wisconsin-Madison
output:
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
    #   toc: true
    #   keep_tex: true
    # pdf_document:
editor_options: 
  chunk_output_type: console
abstract: |
  empty
  
keywords          : "keywords"
wordcount         : "X"
bibliography      : ["r-references.bib"]
floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
#output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences, echo=TRUE}
# selin estimate data
data1 <- read.csv(("/Users/samihabhushan/Downloads/selindatacsv.csv"), header=T)
data1
#library(knitr)
#kable(data1, caption = "Selin Independence Estimates")
```

```{r analysis-preference, echo=TRUE}
# congress sample code
data2 <- read.csv(("/Users/samihabhushan/Downloads/congress.csv"), header=T)
data2
#library(knitr)
#kable(data1, caption = "Congress sample sheet")
```

# Hypothesis
I believe that financial agencies with a higher estimate of agency independence will have lower evidence of congressional influence. Higher rated agencies will have a lower average level of success coded for members of Congress. In turn, those with a lower estimate of independence will have a higher average level of success coded for Congress members, and possibly from a higher level of success coded for members from the same party as the President. 

Further, I believe that Congresspeople that tend to lobby alongside with business entities as opposed to nonprofit institutions will have more asks fulfilled because they are able to engage in more sophisticated lobbying due to additional funding and in turn agencies that fullfill more of these demands will be rated as less independent. 

For each congressperson, success will be coded on a scale from -2 to 1, with -2 being their demand was completely not met and 1 being their demands were fulfilled.

# Theory

I will attempt to answer two questions: first, have federal agencies become increasingly receptive to the demands of members of Congress? Second, are more independent agencies less likely to be influenced by members of Congress? 

I use data from a previous study conducted by Jennifer Selin. Selin's data disseminates various structural features of agencies, such as if the agency is involved within an executive department, if the agency places a conflict of interest restriction on its members, if they provide some avenue for congressional input in the process it nominates its members, and if the agency is a commission or has a board of directors, whether the statute limits the number of members who may serve from the same party. 

With a Bayesian latent variable model, she estimates the level of independence for 321 federal agencies, where the higher the number, the more indendent the agency is in its rulemaking process. Extrapolating data from Selin, we find that financial agencies are generally rated higher in terms of level of independence, where the Consumer Financial Protection Bureau (CFPB) is rated one of the highest at 4.1 and the Federal Trade Commission (FTC) is 2.269. The CFPB's rating is impacted by its 5-year term limit for members, commissioners, and agency heads, additional clauses in the statute that enable members to be removed due to "neglect of duty", and a clause that enables the CFPB to fund a substantial amount of its funding from independent sources. The FTC's rating is higher than most overall agencies because it has similar term limitations and an clause to limits the number of members serving from one party. It has a lower rating than the CFPB likely due to a lower reliance on independent funding sources, therefore making the policy decision process more reliant on government input. 


# Literature Review
Jennifer Selin concludes that "a lack of influence is correlated with statutory provisions that insulate agency policy decisions from review by political principals" (Selin, 2015). This means the main indication of an agency's independence comes from additional clauses in its statute to ensure strict limitations on policy reviews and other decisions from political forces, as opposed to the number of oversight committees for instance. Likewise, the FTC has become more independent over time, according to Selin, due to additional policy features put in place to prevent political interference, including the employment of administrative law judges. Selin found that generally financial agencies had the largest increase in independence over time. 

As stated by King et al. (1994), "the best way to understand a particular event may be by using the methods of scientific inference also to study systematic patterns in similar parallel events." The aforementioned trend should translate beyond the years the study was conducted, translating into the Trump administraton, with agencies such as the FTC and CFPB maintaining or increasing their level of agency independence as much of the structural clauses implemented to maintain agency independence were retained. For example, the U.S. Supreme Court maintained the clause that subject the removal of the director for the CFPB subject to the President's discretion, after a petitioner challenged its constitutionality (Seila Law v. CFPB, 2020). That is to say, while there a high estimate of independence for financial agencies, no agency is immune to political influence. Selin cited that while Securities and Exchange Commission’s (SEC’s) has high estimates, 3.566, SEC policy reflected presidential preferences during the financial crisis in 2008. As such, even with relatively high estimates of independence of financial agencies, we should see some evidence of political influence. 

Wendy Wagner provides comprehensive insight into the phenomenon of "information capture" through a regulatory framework. According to her article, "information capture refers to the excessive use of information and related information costs as a means of gaining control over regulatory decisionmaking in informal
rulemakings." This is often seen in sophisticated lobbying by corporations. For instance, they often use lawyers to use extremely advanced and often confusing language to pressure regulatory agencies into meeting their demands (Wagner, 2010). We can, in turn, see this happen to Congresspeople as well, where lawyers may use language to mislead them into thinking that they are advocating for the people as opposed to a profit-driven incentive. Therefore, more congresspeople that side with businesses not simple for a private purpose, but also because they believe they are serving the people. 

Obtaining information is expensive, and entities with enough funding — often corporations — are able to achieve it. Further, agencies having to process all the information given also overstreches them to beyond their administrative capacity, making them more susceptable to simply giving into the demands. As such, the idea of information capture does play a role in agency independance, where agency that are better able to handle a multitude of information unloaded on them are going to be ranked higher in terms of agency independance. However, an increased presence of congresspeople lobbying alongside businesses may still pressure them into meeting the coalition's demands. 



# Methods
I will gather data from 40 rules issued by financial agencies, including but not limited to the Consumer Financial Protection Bureau, the Treasury Department, and the Office of the Comptroller of the Currency. I focus on public comments issued by members of Congress and focus on whether their demands were ultimately fulfilled in the final rule. 

The research design additionally combines data from a previous study conducted by Jennifer L. Selin, who determined an estimate of the level of structural independence for several agencies during the Obama administration. My study will also include rules during the subsequent Trump administration, wherein the Senate majority switched to the Republican party. I will compare and contrast rules from both eras to determine if influence from political parties have grown over time, including determining if there is an increased presence of politicians commenting on policies. I will also compare the estimates of agency independence as calculated by Selin to level of success with regard to demands made by members of Congress. 

I will also use textual analysis in addition to the hand coding to see if elected officials use more partisan rhetoric over time. In this study, “partisan rhetoric” refers to an individual making positive assertions about their own party or negative assertions about the opposing party. I will also examine whether the demands of elected officials that are the same party as the President are increasingly met over those of the opposing party. 



## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results
I found that there is an increased presence of congresspeople that lobby using more public-facing language. 

# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
