---
title: "BLM Rule Comment Summary"
subtitle: 
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      #toc: true
      # toc_float: true
      code_folding: hide
editor_options: 
  chunk_output_type: console
---


```{r global.options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE, 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)


library(tidyverse)
library(magrittr)
library(tidytext)
library(xml2)
library(knitr)
library(kableExtra)

library(ggplot2); theme_set(theme_bw())
  options(
    ggplot2.continuous.color = "viridis",
    ggplot2.continuous.fill = "viridis"
  )
  scale_color_discrete <- function(...)
    scale_color_viridis_d(..., direction = -1)
  scale_fill_discrete <- function(...)
    scale_fill_viridis_d(..., direction = -1)
  
  
kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "500px")
```

# About   
  
  This file summarizes individual comments for agency rules. To do this it uses three functions; `clean_comments()`, `summarizeText()`, and `summarize_comments()`.  
  
  `clean_comments()` takes in a string of text and cleans it. It is run on each comment provided.  
  
  `summarizeText()` takes in text, max_sentences, custom_stop_words, and agency_name  
  *text* is the string of text for each comment  
  *max_sentences* is a numerical value for the max number of sentences textrank will pull in and summarize  
  *custom_stop_words* is a vector of words to be removed from the dictionary of words in textrank  
  *agency_name* name of the agency to be removed from textrank as a string
function removes custom_stop_words, agency_name, and stop_words from textrank and then uses textrank to rank the sentences by importance.  

`summarize_comments()` takes in text, n_sentences, max_sentences, custom_stop_words, and agency_name  
*n_sentences* is the number of sentences to use for the summary  
function maps `summarizeText()` onto each each comment and then extracts the top *n_sentences*

# Instructions

```{r functions}
# load text cleaning function 
source(here::here("functions", "clean_comments.R"))

# load text summarize function 
source(here::here("functions", "summarizeText.R")) # gives textrank ditionary of words from comments


# load comment summarize function 
source(here::here("functions", "summarize_comments.R")) # ranks sentences and gives top ranked sentence back
```


Change file paths to search for agency and docket

Change custom stop words with agency acronym and rule title

Add agency name to variable
```{r data}
# file paths for comments 
agency <- "BLM"
docket <- "BLM-2013-0002"

# A dataframe of custom stop words, formatted like tidytext::stop_words
# If form letters need to be summariezed standard headings should be removed (docket no., email, etc.)
custom_stop_words <- tibble( 
  words = c("blm", "BLM", "Oil and Gas Hydraulic Fracturing on Federal and Indian Lands")) %>% # ignore case?

  mutate(words = words)%>%
  unnest_tokens(output = word, input = words) %>%
  distinct()

agency_name <- c("bureau of land management")

# get txt file names from a directory, here called “comment_text”
comments <- tibble( path = list.files( here::here('comment_text', agency, docket), 
                                       full.names = T) ) %>% 
  filter( str_detect(path, "txt") )

# in SQL, CFPB file names are regs_dot_gov_document_id, shortened to document_id for now
comments %<>% 
  mutate( document_id = path %>%
            str_remove(".*/")  %>%
            str_remove("-[0-9]*\\..*") 
   ) 


 d <- comments #%>% 
#   # select comments that have been hand coded # No hand coded comments yet
#   filter(document_id %in% coded$document_id)
```

```{r readComment}
read_comment <- . %>%
  read_lines() %>%
  clean_comments() 

# map read
d$text <- map_chr(d$path, read_comment)

```


```{r summarize}
comment_summary <- summarize_comments(d$text, 
                                      n_sentences = 3, 
                                      max_sentences = 50, 
                                      custom_stop_words = custom_stop_words,
                                      agency_name = agency_name)
```


```{r}
d %>% 
  mutate(textrank_summary = comment_summary) %>%
  # a table comparing textrank vs hand coded
  select(document_id, textrank_summary)  %>% kablebox()
```

# Ongoing issues

## Textrank has a hard time with form letters and captures names and titles
```{r}
d %>% 
  mutate(textrank_summary = comment_summary) %>%
  filter(document_id == c("BLM-2013-0002-5231","BLM-2013-0002-0036", "BLM-2013-0002-5667")) %>%
  # a table comparing textrank vs hand coded
  select(document_id, textrank_summary)  %>% kablebox()
```

## Comments with two column formatting in the txt file lead to summaries like this:

```{r}
d %>% 
  mutate(textrank_summary = comment_summary) %>%
  filter(document_id == "BLM-2013-0002-0036") %>%
  # a table comparing textrank vs hand coded
  select(document_id, textrank_summary)  %>% kablebox()
```


## Huh??

```{r}
d %>% 
  mutate(textrank_summary = comment_summary) %>%
  filter(document_id == "BLM-2013-0002-5193") %>%
  # a table comparing textrank vs hand coded
  select(document_id, textrank_summary)  %>% kablebox()
```