---
title: "MasterSql"
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      #toc: true
      #toc_float: true
      #code_folding: show
editor_options: 
  chunk_output_type: console
---


```{r global.options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = FALSE, 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)


library(tidyverse)
library(magrittr)
library(tidytext)
library(xml2)
library(knitr)
library(kableExtra)

library(ggplot2); theme_set(theme_bw())
  options(
    ggplot2.continuous.color = "viridis",
    ggplot2.continuous.fill = "viridis"
  )
  scale_color_discrete <- function(...)
    scale_color_viridis_d(..., direction = -1)
  scale_fill_discrete <- function(...)
    scale_fill_viridis_d(..., direction = -1)
  
  
kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```

# About 

This database contains metadata on over 200,000 dockets and 80,000,000 public comments. It has two tables, `actions` and `comments` corresponding to how documents are organized and identified on regulations.gov (comments are also known as "public submission" documents).

# Instructions

With **R**:

```{r}
library(DBI)
library(RSQLite)

con <- DBI::dbConnect(RSQLite::SQLite(), here::here("db", "master.sqlite"))

dbListTables(con)
```

```{r, eval=FALSE}
action <- dbGetQuery(con, "SELECT * FROM actions")
```

Where the SQL query to get NPRM and Final Rule data for a agency is: 

```{sql, connection=con, output.var = "action"}
SELECT * FROM actions WHERE agency_acronym = 'CFPB'
```

```{r, echo=FALSE}
# rule[1,] %>% mutate_all(as.character) %>% pivot_longer(everything()) %>% kable(format= "pipe")

action %>% kablebox()
```

Likewise, the SQL query to get comments on a docket is

```{sql, connection = con, output.var = "comments", cache=TRUE}
SELECT * FROM comments WHERE docket_ID = 'CFPB-2016-0025'
```

Mutate comment_text is pulling error, 'comment_text' not found
```{r, echo=FALSE}
  comments %>% 
  head() %>%
    #mutate(comment_text = str_sub(comment_text, 1, 140) %>% str_c("...") ) %>% 
    kablebox()
```


```{sql, connection = con, output.var = "action" }

SELECT COUNT(title) FROM actions

```