---
title: "Proposed Rules Open for Comment on `r Sys.Date()`"
#author: "Devin Judge-Lord"
output:
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      #code_folding: hide
editor_options: 
  chunk_output_type: console
---


*Not all documents classified as "Proposed Rules" open for comment on regulations.gov are true proposed rules on rulemaking dockets. Documents open for comment may also include Advanced Notices of Proposed Rulemaking, notices of hearings, and draft Environmental Impact Statements. In my [reasearch](https://judgelord.github.io/research/) on rulemaking, I only use true proposed rules on rulemaking dockets.

Due to API limits, comment counts are based on the most recent 5,000 comments.

```{r global.options, include=FALSE}
source(here::here("setup.R"))


knitr::opts_chunk$set(echo = FALSE, 
                      cache = FALSE)

library(kableExtra)
kablebox <- . %>% 
  head(1000) %>% 
  knitr::kable(digits = 3) %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```


```{r}

# a function to clean raw selected text
clean_summary <- . %>% 
  str_remove_all("./em../mark.|.mark..em.") %>%
  str_replace_all("\\&hellip;\\&nbsp;"," ") %>% 
  str_replace_all("[^[A-z][0-9] \\.\\,\\?\\!\\;&\\;<>]", " ") %>% 
  str_remove_all("\\(|\\)|\\[|\\]") %>%
  str_remove_all("\n") %>% 
  str_squish() %>%
  str_remove_all("`") %>%
  str_squish() %>% str_c("...") %>% str_replace("^", "...")

load(here("data", "open.Rdata"))

ejpr %<>% 
  mutate(EJ = highlightedContent %>% clean_summary() ) %>% 
  select(id, title, openForComment, docketId, frDocNum, commentEndDate, EJ) 

climatepr %<>% 
  mutate(Climate = highlightedContent %>% clean_summary() ) %>% 
  select(id, title, openForComment, docketId, frDocNum, commentEndDate, Climate)

pr_open  %<>% 
  select(id, title, openForComment, docketId, frDocNum, commentEndDate)


open <- full_join(ejpr, climatepr) %>% 
  full_join(pr_open) %>% 
  mutate(Notes = "", Open = openForComment) %>% 
  #filter(comment_start_date > as.Date("2021-01-17")) %>% 
  arrange(rev(as.Date(commentEndDate))) %>% 
  mutate(open_for_comment = as.numeric(openForComment)) %>% 
  arrange(-openForComment)
```

```{r comments, eval=TRUE}
ejcomments %<>% 
    mutate(docketId = str_remove(id, "-[0-9]*$")) %>%
  mutate(highlightedContent = highlightedContent %>% clean_summary() )  %>% 
  select(`EJ Comment` = highlightedContent, docketId, id, title)

climatecomments %<>%
    mutate(docketId = str_remove(id, "-[0-9]*$")) %>%
  mutate(highlightedContent = highlightedContent %>% clean_summary() )  %>% 
  select(`Climate Comment` = highlightedContent, docketId, id, title)

comments <- full_join(ejcomments, 
                      climatecomments
                      )

comments %<>% 
  mutate(URL= str_c("https://www.regulations.gov/document/", id)) %>%
  group_by(docketId, `EJ Comment`, `Climate Comment`) %>%
  summarise(URL = paste(URL, collapse = ";;;") %>% str_remove(";;;.*")) %>% 
  ungroup()

ejcount <- ejcomments %>%  group_by(docketId) %>% 
  count(name = "Unique EJ Comments")

climatecount <- climatecomments %>% group_by(docketId) %>% 
  count(name = "Unique Climate Comments")
```


```{r open, fig.height= 4, fig.width=4.5}
open %<>% mutate(Agency = str_remove(docketId, "(-|_).*"),
                 Contains = case_when(
                   !is.na(Climate) & !is.na(EJ) ~ "Both",
                   !is.na(EJ) ~ "EJ",
                   !is.na(Climate) ~ "Climate",
                   TRUE ~ "Neither"
                 ))

open %>% 
  filter(Open, !is.na(Agency)) %>% 
  ggplot() + 
  aes(x = Agency, fill = Contains) + 
  geom_bar() + 
  coord_flip() + 
  theme_minimal() + 
  labs(y = paste("Proposed Rules Open for Comment\non regulations.gov on", Sys.Date()))

# comments %>% 
#   filter(docketId %in% (open %>% filter(Open) %>% pull(docketId))) %>% 
#   group_by(docketId) %>% 
#   summarise(climate = `Climate Comment` %>% is.na() %>% isFALSE() %>% sum(),
#             ej = `EJ Comment` %>% is.na() %>% isFALSE() %>% sum() ) %>% 
#   pivot_longer(c(ej, climate))
```

```{r o}

# clean up for presentation
o <- open %>% 
  filter(Open) %>% 
    mutate(commentEndDate =commentEndDate %>%  str_remove("T.*") %>% 
                                     replace_na("?"), 
           `Proposed Rule` = str_c("[Comment](https://www.regulations.gov/document/", 
                                   id, ") by ", commentEndDate, 
                                   " on ", docketId, ", ", 
                                   title, " ([Fed. Reg. ", frDocNum, "]( https://www.federalregister.gov/d/", 
                                   frDocNum, "))") ) %>%
    select(docketId, `Proposed Rule`, EJ, Climate #, Open, 
           #`EJ Comments`, `Climate Comments`, 
           #Notes
           ) %>%
  left_join(climatecount) %>% 
  mutate(`Unique Climate Comments` = `Unique Climate Comments` %>% replace_na(0)) %>% 
  left_join(ejcount) %>% 
    mutate(`Unique EJ Comments` = `Unique EJ Comments` %>% replace_na(0))
```

```{r googlesheet, eval= FALSE}
library(googlesheets4)
gs4_auth()
1
1

open %<>% select(Notes, everything())

write_sheet(open, ss = "1Owte4nR18J92xxnPaTDnGgRALZm3-rScCe_co_ohuDU", sheet = "raw")

write_sheet(o, ss = "1Owte4nR18J92xxnPaTDnGgRALZm3-rScCe_co_ohuDU", sheet = "formatted")

write_sheet(comments, ss = "1Owte4nR18J92xxnPaTDnGgRALZm3-rScCe_co_ohuDU", sheet = "comments")
```



### Mentioning "Environmental Justice" and "Climate Change"

```{r}
o %>% drop_na(EJ, Climate) %>% select(-docketId) %>% kablebox()
```

---

### Mentioning "Environmental Justice" but not "Climate Change"

```{r}
o %>% filter(is.na(Climate), !is.na(EJ)) %>% select(-Climate, -docketId) %>% kablebox()
```

---

### Mentioning "Climate Change" but not "Environmental Justice"

```{r}
o %>% filter(is.na(EJ), !is.na(Climate) ) %>% select(-EJ, -docketId) %>% kablebox()
```

---

### Where Comments Raise "Environmental Justice" on Documents that Don't Mention It

```{r}
o  %>% 
    inner_join(comments) %>% 
  filter(is.na(Climate), is.na(EJ), !is.na(`EJ Comment`)) %>% 
  mutate(`EJ Comment` = str_c("[", `EJ Comment`, "](", URL, ")")) %>% 
    select(`Proposed Rule`, `EJ Comment`) %>% 
  kablebox()
```



---

### Where Comments Raise "Climate Change" on Documents that Don't Mention It

```{r}
o  %>% 
    inner_join(comments) %>% 
  filter(is.na(Climate), is.na(EJ), !is.na(`Climate Comment`)) %>% 
  mutate(`Climate Comment` = str_c("[", `Climate Comment`, "](", URL, ")")) %>% 
    select(`Proposed Rule`, `Climate Comment`) %>% 
  kablebox()
```

---

### Raw data [here](https://docs.google.com/spreadsheets/d/1Owte4nR18J92xxnPaTDnGgRALZm3-rScCe_co_ohuDU/edit?usp=sharing) {-}

Code on [github](https://github.com/judgelord/rulemaking)