---
title: "Using the regulations.gov API with **R**"
subtitle: 
author: "Devin Judge-Lord"
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      # code_folding: hide
editor_options: 
  chunk_output_type: console
---


```{r global.options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE, 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)

library(httr)
library(jsonlite)
library(tidyverse)
library(magrittr)
library(knitr)
library(kableExtra)

library(ggplot2); theme_set(theme_bw())
  options(
    ggplot2.continuous.color = "viridis",
    ggplot2.continuous.fill = "viridis"
  )
  scale_color_discrete <- function(...)
    scale_color_viridis_d(..., direction = -1)
  scale_fill_discrete <- function(...)
    scale_fill_viridis_d(..., direction = -1)
  
  
kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```

First, get an API key from https://regulationsgov.github.io/developers/ and Devin's API call functions from https://github.com/judgelord/rulemaking

```{r, cache=FALSE}
source(here::here("functions", "regulations-gov-API-search.R"))

source(here::here("api-key.R"))
```

Each function requires two arguments

1. The `n` most recently posted results (see the "regulations-gov-API-search.R" script for other options for ordering results)

2. Document types (`documenttype`) from this list. 

- "N": Notice, 
- "PR": Proposed Rule, 
- "FR": Rule, 
- "O": Other, 
- "SR": Supporting & Related Material, 
- "PS": Public Submission

To search multiple types, separate them with a comma and a space, e.g. `documenttype = "PR, FR"`.

# Search by document type

```{r}
d <- search.docs(documenttype = "PR" , 
                 n = 1000)

head(d) %>% select(title, everything() ) %>% kablebox()

d$postedDate %<>% as.Date()
d$commentStartDate %<>% as.Date()

ggplot(d) + 
  aes(x = postedDate, fill = documentType) +
  geom_histogram()
```


# Search by agency

Select agencies, again seperating multiple agencies with a comma and a space

```{r}
d <- search.agency.docs(agency = "SEC, FAA, FRA", 
                        documenttype= "PR, FR",
                        n = 2000)

head(d) %>% select(title, everything() ) %>% kablebox()

d$postedDate %<>% as.Date()
d$commentStartDate %<>% as.Date()

# plot histogram by by agency over time
ggplot(d) + 
  aes(x = postedDate, fill = documentType) +
  geom_histogram() + 
  facet_grid(rows = "agencyAcronym", scales = "free_y")
```


# Search by docket

```{r}
# Select a docket
docket <- "EPA-HQ-OAR-2017-0355"

# Search
d <- search.docket(docket = "EPA-HQ-OAR-2017-0355", 
                   documenttype = "PS",
                   n = 2000)

head(d) %>% select(title, everything() ) %>% kablebox()

d$postedDate %<>% as.Date()
d$commentStartDate %<>% as.Date()

ggplot(d) + 
  aes(x = postedDate, fill = documentType) + 
  geom_histogram() + 
  labs(title = docket)
```

# Search by keyword(s)

> Note: keyword serches return an additional feature "summary" which is text before and after the keyword in the comment text

```{r}
keywords = "environmental justice"

d <- search.keywords(keywords = c("environmental justice"),
                     documenttype = "PS",
                     n = 2000)

head(d) %>% select(title, everything() ) %>% kablebox()

d$postedDate %<>% as.Date()
d$commentStartDate %<>% as.Date()

ggplot(d) + 
  aes(x = postedDate, fill = documentType) + 
  geom_histogram() + 
  labs(title = str_c("Documents containing \"", keywords, "\""))
```






