---
title: "Rules summarized with `textrank`"
subtitle: 
author: ""
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
editor_options: 
  chunk_output_type: console
---


```{r global.options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE, 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)


library(tidyverse)
library(magrittr)
library(tidytext)
library(xml2)
library(knitr)
library(kableExtra)

library(ggplot2); theme_set(theme_bw())
  options(
    ggplot2.continuous.color = "viridis",
    ggplot2.continuous.fill = "viridis"
  )
  scale_color_discrete <- function(...)
    scale_color_viridis_d(..., direction = -1)
  scale_fill_discrete <- function(...)
    scale_fill_viridis_d(..., direction = -1)
  
  
kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```

`summarize_sections` is a wrapper for `textrank_sentences` that first splits a rule into sections and then returns a summary for each. It takes three inputs:

1. Rule `text` extracted from a Federal Register xml file with the `xml_rule_text()` function (or any other data frame with a column called "text").
1. The number of sentences to return, `n_sentences` (optional)--this controls the length of the summary. Do you want each section summarized in one sentence summary or five? The current default is two sentences per rule section, but I plan to make this vary by a section's length (longer sections may deserve a longer summary).  
1. The maximum number of sentences per section to summarize `max_sentences` (optional). Fewer sentences makes textrank go faster, but we will often want to use all sentences. The default is 100 sentences per rule section.

You can download it [here](https://github.com/judgelord/rulemaking/blob/master/functions/textrank.R) or load it in R with `source("https://raw.githubusercontent.com/judgelord/rulemaking/master/functions/textrank.R")`

For more on textrank and the steps involved in applying it to a rule, see the PDF that Chris shared. 

```{r}
# load summarize_sections() function
source("https://raw.githubusercontent.com/judgelord/rulemaking/master/functions/textrank.R")
```

# The Volcker Rule 

### Proposed Rule
```{r}
xml_rule_text("https://www.federalregister.gov/documents/full_text/xml/2011/11/07/2011-27184.xml") %>% 
  summarize_sections(n_sentences = 2,
                     max_sentences = 100) %>% 
  kablebox()
```

### Final Rule
```{r}
xml_rule_text("https://www.federalregister.gov/documents/full_text/xml/2014/01/31/2013-31511.xml") %>% 
  summarize_sections() %>% 
  kablebox()
```

# CFPB's Payday Loan Rule

### 2016 Proposed Rule
```{r}
xml_rule_text("https://www.federalregister.gov/documents/full_text/xml/2016/07/22/2016-13490.xml") %>% 
  summarize_sections() %>% 
  kablebox()
```

### 2017 Final Rule
```{r}
xml_rule_text("https://www.federalregister.gov/documents/full_text/xml/2017/11/17/2017-21808.xml") %>% 
  summarize_sections() %>% 
  kablebox()
```

### 2019 Proposed Rule
```{r}
xml_rule_text("https://www.federalregister.gov/documents/full_text/xml/2019/02/14/2019-01906.xml") %>% 
  summarize_sections() %>% 
  kablebox()
```

### 2020 Final Rule
```{r}
xml_rule_text("https://www.federalregister.gov/documents/full_text/xml/2020/07/22/2020-14935.xml") %>% 
  summarize_sections() %>% 
  kablebox()
```
