---
title: "Rules summarized with `textrank`"
subtitle: 
author: ""
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
editor_options: 
  chunk_output_type: console
---


```{r global.options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE, 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)


library(tidyverse)
library(magrittr)
library(tidytext)
library(xml2)
library(knitr)
library(kableExtra)

library(ggplot2); theme_set(theme_bw())
  options(
    ggplot2.continuous.color = "viridis",
    ggplot2.continuous.fill = "viridis"
  )
  scale_color_discrete <- function(...)
    scale_color_viridis_d(..., direction = -1)
  scale_fill_discrete <- function(...)
    scale_fill_viridis_d(..., direction = -1)
  
  
kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```

`summarize_comments` is a wrapper for `textrank_sentences` that returns a summary for each comment. It takes three inputs:

1. Comment `text` 
1. The number of sentences to return, `n_sentences` (optional)--this controls the length of the summary. Do you want each comment summarized in one sentence summary or five? The current default is two sentences per comment, but I plan to make this vary by a comment's length (longer sections may deserve a longer summary).  
1. The maximum number of sentences per comment to summarize `max_sentences` (optional). Fewer sentences makes textrank go faster, but we will often want to use all sentences. The default is 100 sentences per comment.

When the function is done, I'll add it to the textrank.R file.

You can download it [here](https://github.com/judgelord/rulemaking/blob/master/functions/textrank.R) or load it in R with `source("https://raw.githubusercontent.com/judgelord/rulemaking/master/functions/textrank.R")`


```{r}
# load summarize_sections() function
source("https://raw.githubusercontent.com/judgelord/rulemaking/master/functions/textrank.R")
```

# CFPB Payday Loan Rule

```{r summarize}

# text %>%
#   filter(document_id == "CFPB-2016-0025-1748-1") %>%
#   select(text)


summarizeText <- function(text) {
  
  sentences <- commentText %>%
    select(text) %>%
    unnest_sentences(output = sentences, input = text) %>%
    distinct() %>%
    mutate(textrank_id = row_number()) %>% 
    # textrank requires columns in order
    select(textrank_id, sentences)

  # select max sentences per section
  sentences %<>% filter(textrank_id <= 400)

  words <- unnest_tokens(sentences, output = word, input = 'sentences') %>% 
    distinct() %>% 
    anti_join(tidytext::stop_words)
  
  if(nrow(sentences) > 1){
out <- textrank::textrank_sentences(data = sentences, 
                                    terminology = words)
  } else {
    out <- list(sentences = as.data.frame(sentences))
  }

return(out)

}

summarizeText(commentText)
#summary %>%   knitr::kable()
```

