---
title: "The Volcker Rule"
subtitle: 
author: ""
output:
    # pdf_document:
    #   toc: true
    #   keep_tex: true
    html_document:
      highlight: zenburn
      toc: true
      toc_float: true
      code_folding: hide
editor_options: 
  chunk_output_type: console
---


```{r global.options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = FALSE, 
                      fig.width=8.5, 
                      split = T,
                      fig.align = 'center', 
                      fig.path='figs/',
                      warning=FALSE, 
                      message=FALSE)


library(tidyverse)
library(magrittr)
library(tidytext)
library(xml2)
library(knitr)
library(kableExtra)

library(ggplot2); theme_set(theme_bw())
  options(
    ggplot2.continuous.color = "viridis",
    ggplot2.continuous.fill = "viridis"
  )
  scale_color_discrete <- function(...)
    scale_color_viridis_d(..., direction = -1)
  scale_fill_discrete <- function(...)
    scale_fill_viridis_d(..., direction = -1)
  
  
kablebox <- . %>%  knitr::kable() %>% 
  kable_styling() %>% 
  scroll_box(height = "400px")
```


```{r}
# links to proposed and final rules
pr <- "https://www.federalregister.gov/documents/full_text/xml/2011/11/07/2011-27184.xml"
fr <- "https://www.federalregister.gov/documents/full_text/xml/2014/01/31/2013-31511.xml"

# text marking the end of the preamble
rule_text_starts <- . %>% {which(str_detect(., "The text of the proposed common rules|^Text of Common Rule"))}

# a function to extract text
xml_rule_text <- . %>% 
  read_xml() %>% 
  xml_children() %>% 
  xml_children() %>% 
  xml_text() %>% 
  str_squish() %>% 
  tibble::enframe(value = "text",  name = "id") 
  
clean <- . %>%  
  filter(nchar(text) > 60) %>% 
  mutate(id = str_pad(id, 4, pad = "0") %>% 
           paste(text) %>% 
           str_sub(1,80) %>%
           str_c("...")) %>% 
  # remove numbering 
  mutate(text = text %>% str_replace_all("\\(.\\)|[0-9]|_", " "))
```

# Proposed Rule
```{r}
# extract proposed rule text
pr %<>% xml_rule_text() 

pr %<>%
  mutate(preamble = id < rule_text_starts(pr$text))  %>%
  clean()

# preamble
# pr %>% kablebox()

# rule text 
pr %>% filter(preamble == F) %>% kablebox()
```

# Final Rule
```{r}
fr %<>% xml_rule_text() 

fr %<>%
  mutate(preamble = id < rule_text_starts(fr$text)) %>% 
  clean

# fr preamble text 
# fr %>% kablebox()

# rule text 
fr %>% filter(preamble == F) %>% kablebox()

# tengrams
pr10 <- pr %>% unnest_tokens(tengram, text, token = "ngrams", n = 10) %>% filter(!is.na(tengram))

fr10 <- fr %>% unnest_tokens(tengram, text, token = "ngrams", n = 10) %>% filter(!is.na(tengram))
```

# What changed? 

A straightforward method is a 10-gram moving window; detecting 10-word (or longer) phrases from a Proposed Rule are left unchanged in the Final Rule. 10-word phrases from the Proposed Rule that do not appear Final Rule may have been cut or merely changed slightly. Likewise, 10-word phrases from the Final Rule that do not appear in the Proposed Rule may reflect minor edits or entirely new text.

Pros:  
- straightforward and easy to explain  
- can be extremely sensitive to moderately sensitive depending on how texts are cleaned  
- computationally cheap

Cons:  
- less precise than more computationally-intense methods

## Proposed rule text that was cut or changed in the Final Rule
```{r pr_tengrams, fig.height=9}
# rule text without preamble
frr10 <- filter(fr10, !preamble)
prr10 <- filter(pr10, !preamble)

prr10 %<>% 
  mutate(new = !tengram %in% frr10$tengram,
         word1 = str_extract(tengram, "\\w+") )

prr10 %<>% 
  group_by(id) %>%
  # mark transitions from new to new or new to new text
  mutate(trans = ifelse(new & !lead(new) | !new & lead(new), T, F)) %>% 
  # if old, the next 9 are also old 
  mutate(fix =  reduce(map(1:9, ~ lag(trans, ., 0)), `+`)) %>%
  mutate(new = ifelse(fix>0 & new == T, F, new)) %>% 
  # add elipses after transition words 
  mutate(word1 = ifelse(trans, str_c(word1, "..."), word1)) 

prr10 %<>% 
  # collapse words
  group_by(new, id) %>%
  summarise(word_count = n(),
            text = paste(word1, collapse = " ")) 

# counts per section 
prr10 %>% 
  select(-text) %>% 
  spread(key = new, value = word_count) %>% 
  rename(`Words from phrases that were also in the Final Rule` = `FALSE`, 
         `Words not from phrases that were also in the Final Rule` = `TRUE`)  %>% 
  kablebox()

prr10 %>% 
  ggplot() +
  aes(x = id, y = word_count, fill = new) + 
  geom_col() +
  coord_flip() + 
  labs(fill = "Cut or Changed in Final Rule", 
       x = "",
       "Number of words") + 
  facet_wrap("id", scales = "free", ncol = 1)  +
  theme_minimal() + 
  theme(axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        strip.text = element_text(hjust = 0),
        legend.position =  "top")  

# look at a section
prr10 %>% 
  #slice(1) %>% 
  ungroup() %>% 
  select(-word_count) %>% 
  spread(key = new, value = text) %>% 
  rename(`Words from phrases that were also in the Final Rule` = `FALSE`, 
         `Words not from phrases that were also in the Final Rule` = `TRUE`)  %>%
  kablebox()
```

---

## Final Rule text that was added or changed from the Proposed Rule
```{r fr_tengrams, fig.height=15}
# rule text without preamble
frr10 <- filter(fr10, !preamble)
prr10 <- filter(pr10, !preamble)

# final rule 10-grams not in proposed rule 
frr10 %<>% 
  mutate(new = !tengram %in% prr10$tengram,
         word1 = str_extract(tengram, "\\w+") )

frr10 %<>% 
  group_by(id) %>%
  # mark transitions from new to new or new to new text
  mutate(trans = ifelse(new & !lead(new) | !new & lead(new), T, F)) %>% 
  # if old, the next 9 are also old 
  mutate(fix =  reduce(map(1:9, ~ lag(trans, ., 0)), `+`)) %>%
  mutate(new = ifelse(fix>0 & new == T, F, new)) %>% 
  mutate(word1 = ifelse(trans, str_c(word1, "..."), word1)) 

frr10 %<>% 
  # collapse words
  group_by(new, id) %>%
  summarise(word_count = n(),
            text = paste(word1, collapse = " ")) 

frr10 %>% 
  ggplot() +
  aes(x = id, y = word_count, fill = new) + 
  geom_col() +
  coord_flip() + 
  labs(fill = "Added or Changed in Final Rule", 
       x = "",
       "Number of words") + 
  facet_wrap("id", scales = "free", ncol = 1)  +
  theme_minimal() + 
  theme(axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        strip.text = element_text(hjust = 0),
        legend.position =  "top")  

frr10 %>% 
  ungroup() %>% 
  select(-word_count) %>% 
  spread(key = new, value = text) %>% 
  rename(`Words from phrases that were also in the Proposed Rule` = `FALSE`, 
         `Words not from phrases that were also in the Proposed Rule` = `TRUE`)  %>%
  kablebox()
```

<!---

## Final rule **preamble** text that was not in the Proposed Rule preamble

```{r prp_tengrams, eval = TRUE, fig.height=8}
# preamble text
frp10 <- filter(fr10, preamble)
prp10 <- filter(pr10 %>% head(), preamble)

frp10 %<>% 
  mutate(new = tengram %in% prp10$tengram,
         word1 = str_extract(tengram, "\\w+") )

frp10 %<>% 
  # mark transitions from new to new or new to new text
  mutate(trans = ifelse(new & !lead(new) | !new & lead(new), T, F)) %>% 
  # if FALSE, the next 9 are also FALSE 
  mutate(fix =  reduce(map(1:9, ~ lag(trans, ., 0)), `+`)) %>%
  mutate(new = ifelse(fix>0 & new == T, F, new)) %>% 
  mutate(word1 = ifelse(trans, str_c(word1, "..."), word1)) %>% 
  # collapse words
  group_by(new, id) %>%
  summarise(word_count = n(),
            text = paste(word1, collapse = " "))

# counts per section 
frp10 %>% 
  select(-text) %>% 
  spread(key = new, value = word_count) %>% 
  rename(`In Final Rule` = `FALSE`, 
         `Not in Final Rule` = `TRUE`)  %>% 
  kablebox()

frp10 %>% 
  .[1:12,] %>% 
  ggplot() +
  aes(x = id, y = word_count, fill = new) + 
  geom_col() +
  coord_flip() + 
  labs(fill = "Changed in Final Rule", x = "",
       y = "Number of words",
       title = "(plot only covers the first 10% of preamble subsections)") + 
  facet_wrap("id", scales = "free", ncol = 1)  +
  theme_minimal() + 
  theme(axis.text.y = element_blank(),
        panel.grid.major.y = element_blank(),
        strip.text = element_text(hjust = 0),
        legend.position =  "top")

# look at a section
frp10 %>% 
  ungroup() %>% 
  select(-word_count) %>% 
  spread(key = new, value = text) %>% 
  rename(`Words from Proposed Rule Preamble` = `FALSE`, 
         `Words not from Proposed Rule Preamble` = `TRUE`)  %>% 
  kablebox()
```

-->